Application of computational fluid dynamics and process modelling to investigate low-load operation of a subcritical utility-scale boiler.

### 2021-11-05 PAPER_2/PAPER_2_REV1.tex
Update - 05/11/2021

\journal{Energy}

@@

\title{Numerical investigation of low-load operation of a subcritical power boiler using CFD and process modelling}

%% Group authors per affiliation:
\author{B.T. Rawlins}
\author{R. Laubscher\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{rlaubscher@sun.ac.za}
\author{P. Rousseau}
\address{Department of Mechanical Engineering, Applied Thermal-Fluid Process Modelling Research Unit, University of Cape Town, Library Rd, Rondebosch, Cape Town, 7701, South Africa}
\address{$^*$ Department of Mechanical Engineering, Stellenbosch University, Stellenbosch Central, Stellenbosch, 7599, South-Africa}

\begin{abstract}
Low-load operation of a 620 $MWe$ sub-critical boiler is analysed with the aid of a co-simulation methodology that incorporates a detail three dimensional computational fluid dynamics (CFD) model together with a one dimensional process model. A discretized one dimensional model of the water/steam circuit was developed for the furnace, the radiant superheaters and the convective/back pass heat exchangers. This is coupled with a detailed CFD model incorporating the furnace and radiant superheaters using a Eulerian-Eulerian reference frame. The coupled model is used to investigate the combustion stability, water/steam side effects, and the radiant heat exchanger operational limits for six different burner firing arrangements at a low boiler load of 32\% of Maximum Continuous Rating. The results show that certain firing arrangements can lead to a high risk of fire-side corrosion and overheating of heat exchanger components. Based on analyses of combustion stability, boiler efficiency and the safe operation of heat exchanger components, a mixed firing arrangement with a higher secondary air mass flowrate for non-firing burners was selected as the best operational strategy at this low load for the boiler under investigation.
\end{abstract}

@@

%\linenumbers

@@

Using the process model of Figure \ref{fig_flownex} and the results of the CFD simulations the important process control parameters were determined. Table \ref{tbl_process_parameters} summarizes the results with the highest boiler efficiencies observed for Cases 3 and 6. With a lower SA flowrate Cases 4 to 6 exhibit a higher boiler efficiency due to the decrease in dry gas losses, when compared to Cases 1 to 3. All cases exhibit adequate control of the main steam exit temperature by use of ATT1 and ATT2. The exit temperature of the reheaters, for all Cases, are determined to within the 20 [$^\circ C$] tolerance for the intermediate turbine inlet conditions, as stipulated in the design C-schedules for the plant. However, with the lack of ATT-RH control, a sudden decrease in steam generation can lead to RH overheating and possible RH failure. Considering all of the above, Cases 3 and 6 are seen as the best firing arrangements for continuous low-load operation at a load of 32\% MCR.\\

@@

\clearpage
\section{Conclusions}
The present study demonstrates the use of a coupled CFD/process model to study the effects of burner configuration at a boiler load of 32\% MCR. A validation case was first performed for the CFD model, which showed adequate results for heat fluxes to the walls, with a general tendency of the model to under-predict the heat fluxes to the furnace. The model also showed sufficient accuracy in determining the combustion characteristics at the 100 \% MCR load. \\

>>>>

\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\centerline{\thepage}}%
 \let\@evenfoot\@oddfoot}
\makeatother

@@

\title{Application of computational fluid dynamics and process modelling to investigate low-load operation of a subcritical utility-scale boiler.}

%% Group authors per affiliation:
\author{B.T. Rawlins $^{a}$ \corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{rwlbra001@myuct.ac.za}
\author{R. Laubscher $^b$}
\author{P. Rousseau $^a$}
\address{$^a$ Department of Mechanical Engineering, Applied Thermal-Fluid Process Modelling Research Unit, University of Cape Town, Library Rd, Rondebosch, Cape Town, 7701, South Africa}
\address{$^b$ Department of Mechanical Engineering, Stellenbosch University, Stellenbosch Central, Stellenbosch, 7599, South-Africa}

\begin{abstract}
To enable the inclusion of intermittent renewable energy sources on existing power grids requires base-load coal-fired power plants to operate flexibly, including fast load changes and continuous low-load operation. Low-load operation of a 620 $MWe$ sub-critical boiler is analysed with the aid of a co-simulation methodology that incorporates a detail three dimensional computational fluid dynamics (CFD) model together with a one dimensional process model. A discretized one dimensional model of the water/steam circuit was developed for the furnace, the radiant superheaters and the convective/back pass heat exchangers. This is coupled with a detailed CFD model incorporating the furnace and radiant superheaters using a Eulerian-Eulerian reference frame. The coupled model is used to investigate the combustion stability, water/steam side effects, and the radiant heat exchanger operational limits for six different burner firing arrangements at a low boiler load of 32\% of Maximum Continuous Rating. The results show that certain firing arrangements can lead to a high risk of fire-side corrosion and overheating of heat exchanger components. Based on analyses of combustion stability, boiler efficiency and the safe operation of heat exchanger components, a mixed firing arrangement with a higher secondary air mass flowrate for non-firing burners was selected as the best operational strategy at this low load for the boiler under investigation.
\end{abstract}

@@

\linenumbers

@@

Using the process model of Figure \ref{fig_flownex} and the results of the CFD simulations the important process control parameters were determined. Table \ref{tbl_process_parameters} summarizes the results with the highest boiler efficiencies observed for Cases 3 and 6. With a lower SA flowrate Cases 4 to 6 exhibit a higher boiler efficiency due to the decrease in dry gas losses, when compared to Cases 1 to 3. All cases exhibit adequate control of the main steam exit temperature by use of ATT1 and ATT2. The exit temperature of the reheaters, for all Cases, are determined to within the 20 [$^\circ C$] tolerance for the intermediate turbine inlet conditions, as stipulated in the design C-schedules for the plant. However, with the lack of ATT-RH control, a sudden decrease in steam generation can lead to RH overheating and possible RH failure. Considering all of the above, Cases 3 and 6 are seen as the best firing arrangements for continuous low-load operation at a load of 32\% MCR.

@@

\section{Conclusions}
The present study demonstrates the use of a coupled CFD/process model to study the effects of burner configuration at a boiler load of 32\% MCR. A validation case was first performed for the CFD model, which showed adequate results for heat fluxes to the walls, with a general tendency of the model to under-predict the heat fluxes to the furnace. The model also showed sufficient accuracy in determining the combustion characteristics at the 100 \% MCR load. \\
